---
title: "simpleRegress"
author: "aubrey garcia"
date: "summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

For model length~neck, adjusted $r^2$ = 0.75


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


##length based on age

```{r}
agemodel<- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(agemodel)
summary.lm(agemodel)
```

adjust $r^2$ for the age model is 0.51, this is less than the adjusted $r^2$ for neck so this model is not as good as the neck model. 

###Length bases on chest

```{r}
chestmodel<- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestmodel)
summary.lm(chestmodel)
```
the adjusted $r^2$ for chest =0.79, which is better than the adjusted $r^2$ = 0.75 for neck.


### Length Based on headwth

```{r}
headwthmodel<- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwthmodel)
summary.lm(headwthmodel)
```
the adjusted $r^2$ for HEADWTH is 0.54 which means it is not better than the adjusted $r^2$ for chest =0.79

###Length bases on sex
```{r}
sexmodel<- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexmodel)
```

The adjusted $r^2$ for SEX is 0.01, which is way worse the the adjusted $r^2$ for chest =0.79

###Length based on HEADLEN


